# Preface {#preface .unnumbered}

```{r 'setup', include=FALSE}
source('_common.R')
```

## Guidelines

- The project will be kept in sync with the latest versions of both Python and R. 
- It will be ensured that the code can be executed by direct copy-paste and without significantly affecting user work environment
- Equality & Identity tests related to language rules may be hidden to avoid clutter

## References

- [CRAN R Manuals](https://cran.r-project.org/doc/manuals/) or [CRAN Duplicate](https://cran.r-project.org/manuals.html) :
  - [An Introduction to R](https://cran.r-project.org/doc/manuals/R-intro.html)
  - [R Installation and Administration](https://cran.r-project.org/doc/manuals/R-admin.html)
  - [R Data Import/Export](https://cran.r-project.org/doc/manuals/R-data.html)
  - [Writing R Extensions](https://cran.r-project.org/doc/manuals/R-exts.html)
  - [R Internals](https://cran.r-project.org/doc/manuals/R-ints.html)
  - [R Language Definition](https://cran.r-project.org/doc/manuals/R-lang.html)
  - [R FAQ](https://cran.r-project.org/doc/manuals/R-FAQ.html)

- [Python](https://docs.python.org/3/index.html) :
  - [Python Standard Library](https://docs.python.org/3/library/index.html)
  - [Python Language Reference](https://docs.python.org/3/reference/index.html)
  - [Python Tutorial](https://docs.python.org/3/tutorial/index.html)
  - [Python HOWTO](https://docs.python.org/3/howto/index.html)


## Prerequisites

```{r 'R-Version', decorate=TRUE}
strsplit(R.version.string, ' ')[[1]][3]           # R Version
```

```{python 'Y-Version', decorate=TRUE}
print(sys.version.split()[0])                     # Python Version

```

## Environment

```{r 'R-Attached-Packages', decorate=TRUE}
# Attached (Basic) Packages in R 
sessionInfo()
```

```{python 'Y-Loaded-Modules', decorate=TRUE}
# Imported (Basic) Modules in Python 
allmodules = [v.__name__ for k, v in globals().items() 
    if type(v) is types.ModuleType and not k.startswith('__')]

len(allmodules)                                   # Count the Loaded Modules

', '.join(allmodules)                             # List Modules

```

```{r 'R-Python-Configuration', decorate=TRUE}
# Python Configuration with R Package 'Reticulate'
py_config()
```

