# Introduction {#introduction-p01}

```{r 'setup', include=FALSE}
################################################################################
source("_common.R")
if(TRUE) c("lobstr") |> 
  lapply(library, character.only=TRUE)
```

## Basics

- Both R and Python are case-sensitive
- Both R and Python use the Hash sign `#` to comment out anything after it, till the newline
- Both R and Python keep the Backslash `\` as reserved to escape the character that follows it

## Mathematical Operators

- R and Python both have following operators:  
  - Addition `+`, Subtraction `-`, Multiplication `*`, Division `/`
  - Exponent is given by caret `^` or `**` in R and by `**` in Python
    - Python uses caret for bitwise `XOR`
  - Modulus (Remainder) is given by `%%` in R and by `%` in Python
    - Python uses `%%` to escape the percent sign in strings
    - R does not use standalone `%` as an operator (except for special cases like `sprintf()`)
    - However, R uses `%...%` syntax to define some special operators (infix operators) like `%*%`, `%in%`, etc.
  - Integer Division (Quotient) is given by `%/%` in R and by `//` in Python
    - Python does not use `%/%` and R does not use `//`

```{r 'R-Operators-Math', decorate=TRUE}
# Exponent
5 ** 3
5 ^ 3

5 %% 3    # Modulus (Remainder)
5 %/% 3   # Integer Division (Quotient)
```

```{python 'Y-Operators-Math', decorate=TRUE}
# Exponent
5 ** 3

5 % 3     # Modulus (Remainder)
5 // 3    # Integer Division (Quotient)

```

## Help

- Both R and Python have `help()` to get details about the topic
  - R also has a shorthand to get help by prefixing `?` before the topic

```{r 'R-Help', decorate=TRUE, eval=FALSE}
if(FALSE) {
  help(print)
  ?print
  ?`+`
}
```

```{python 'Y-Help', decorate=TRUE, eval=FALSE}
if(False):
    help(print)

```

## Print

- Printing is largely handled by `print()` in both R and Python
  - Explicit and Implicit both types of printing is available


## Braces & Indentation

- R uses braces `{}` to group expressions in code blocks and ignores indentation
- Indentation and white spaces are extremely significant in Python (unlike R)
  - 4 Spaces (not Tab) are being used, in this book, for Python chunks
  - A backslash `\` allows you to break up one long piece of code into several parts
  - Any character (including whitespace) after the backslash will cause an error
  - Parentheses `()` can be used to enclose all of the code. Parentheses work like the backslash, but allow extra characters at the end
  - An extra newline is being added in Python chunks, in this book, to indicate that the command /block has been completed

## NA NaN 0/0 - nan

- 0/0 is `NaN` (Not a Number) in R but is undefined in Python
  - The IEEE754 standard defines three special floating-point numbers, `Inf`, `-Inf` and `NaN`
  - `NaN` is 'not well-defined real number'
  - In numerical computing, if the result can not be written in regular real number, or `Inf` or `-Inf`, it is expressed as `NaN`. 
- `NA` and `NaN`
  - Python does not distinguish between 'missing values' and 'not a number'. whereas, R clearly distinguish between `NA` and `NaN`
  - `NaN` is a float value, so we cannot have explicit missing values in non-float columns. In R, `NA` can be of any type.

- Comparison operation with `NaN` (even with itself) 
  - Python: `==` returns `False`, `!=` returns `True` (IEEE754 Standard)
  - R: All such operations (`==`, `!=`, `>=` `<=`, `>`, `<`) results in `NA` 
  - R: `NA` or `NaN` compared to anything is `NA`.

- Coercion of `NaN` in R (Or Casting or Conversion)
  - `NaN` coercion to `integer` gives `NA_integer_` 
  - `NaN` coercion to `logical` gives `NA`
  - `NaN` coercion to `numeric` gives `NaN` 
  - `NaN` coercion to `character` gives `"NaN"`
  
- Reticulate: 
  - In python, missing values are also `NaN`
  - Both `np.nan` and `math.nan` of Python are converted to `NaN` of R
  - `NaN` of R is converted to `NaN` of Python 
  
- Pandas
  - Both `np.nan` and `math.nan` are treated as 'missing values' with class as `numpy.float64`
  - `None` is also treated as a missing value. For numeric columns, `None` is converted to `nan`. For columns of `object` type, it is kept as `None` but is treated as missing value.
  - There exist a `pd.NA` scalar in Pandas, similar to R `NA`. However, it is experimental and currently treated as `np.nan` internally. So, not being used for now, in this book.

```{r 'R-0by0', decorate=TRUE}
# 0/0 is defined as NaN in R (unlike Python)
is.nan(NaN)
is.nan(0/0)

# NaN is treated as NA, however, NA is not treated as NaN
is.na(NaN)
is.nan(NA)

# NaN is numeric
stopifnot(is.numeric(NaN))

# Comparison with NaN: All operations are NA, use is.nan() (unlike IEEE, Python)
stopifnot(all(identical(NaN == NaN, NA),
              identical(NaN != NaN, NA),
              identical(NaN <= NaN, NA),
              identical(NaN >= NaN, NA),
              identical(NaN <  NaN, NA),
              identical(NaN >  NaN, NA)))

# Coercion of NaN
stopifnot(all(identical(as.integer(NaN), NA_integer_),
              identical(as.logical(NaN), NA),
              identical(as.character(NaN), "NaN"),
              identical(as.numeric(NaN), NaN)))
```

```{python 'Y-0by0', decorate=TRUE}
# 0/0 is undefined in Python (unlike R)
try:
    print(0/0)
except ZeroDivisionError as e:
    print(e) #e.args type(e)

# There are Two NaN in modules NumPy and Math
# Use module specific methods but not the 'is' keyword
pp = float("NaN") # Not case sensitive i.e. NaN, NAN, nan etc. all are valid
print(pp)
type(pp)

assert(math.isnan(math.nan) and math.isnan(np.nan) and math.isnan(pp))
assert(np.isnan(np.nan) and np.isnan(math.nan) and np.isnan(pp))

# 'is' returns True if and only if the two references are to the same object
np.nan is math.nan            # False

# Comparison with NaN: == is False, != is True (as per IEEE but unlike R)
np.nan == np.nan
math.nan == math.nan
 
np.nan != np.nan
math.nan != math.nan

```

## Boolean

- R (TRUE, FALSE) vs. Python (True, False) - Case is different
- Python:
  - All values are considered "truthy" except for the following, which are "falsy"
    - `None`, `False`, `0`, `0.0`, `0j`, `Decimal(0)`, `Fraction(0, 1)`
    - Empty: list `[]`, dictionary `{}`, tuple `()`, string `''`, range `range(0)` etc.
    - objects for which:
      - `obj.__bool__()` returns `False`
      - `obj.__len__()` returns `0`
    - A "truthy" value will satisfy the check performed by `if` or `while` statements. 
    - We use "truthy" and "falsy" to differentiate from the bool values `True` and `False`.
    - Python determines the truthiness by applying `bool()` to the type, which returns `True` or `False` which is used in an expression like `if` or `while`. 
      - Instances of a user-defined class are considered truthy by default.


## Short-circuit

- Both R and Python short-circuit the evaluation
  - However, in Python, `and` and `or` are lazy whereas `&` and `|` are not

```{r 'R-Short-Circuit', decorate=TRUE}
# Only the First expression is evaluated if it is enough to determine the result
TRUE | NaN
FALSE & NaN

# Here the Second expression is also evaluated because first is not sufficient
TRUE & NaN
FALSE | NaN

# Logical evaluation returns TRUE, FALSE or NA in R (unlike Python)
# 0 is FALSE
3 | NaN
0 & NaN
0 | NaN
```

```{python 'Y-Short-Circuit', decorate=TRUE}
# Only the First expression is evaluated if it is enough to determine the result
True or np.nan
False and np.nan

# Here the Second expression is also evaluated because first is not sufficient
True and np.nan 
False or np.nan

# Logical evaluation returns the value of final expression evaluated (unlike R)
3 or np.nan
True and 3

```

## NULL - None

- `NULL` represents the null object in R: it is a reserved word. NULL is often returned by expressions and functions whose value is undefined. 
  - There is only one null object in memory
  - Inside if condition, `NULL` throws error, use `is.null(x)` syntax
  - If the output of an expression does not show numbers in brackets like '[1]' then it is a NULL type return. [Numbers] show that it is a Vector. Ex: `str()` and `cat()` outputs are of NULL Type.
- Python uses the keyword `None` to define null objects and variables
  - Inside `if` condition, `None` acts as `False`, use `x is None` syntax
  - `None` is a singleton. There is only one `None` in memory
  - [Real Python](https://realpython.com/null-in-python/)

```{r 'R-NULL', decorate=TRUE}
# Return of str() is NULL which is passed to print()
print(str("This returns NULL in R"))

# Assign NULL
aa <- NULL
# Type
typeof(aa)
# Check
is.null(aa)
# is.na() returns logical(0) for NULL
is.na(aa)

# A name pointing to NULL is different from a name which does not exist 'bb'
exists("aa")
exists("bb")
tryCatch(expr = print(bb), error = \(e) print(e))

# Usage inside if conditional: NULL throws Error, use is.null()
if(is.null(aa)) {
  print("Object is NULL") 
} else {
  print("Object is Not NULL")
}
```

```{python 'Y-None', decorate=TRUE}
# Return of print() is None which is passed to print()
print(print("This returns None in Python"))

# Assign None
pp = None
# Type
type(pp)
# Check
pp is None

# A name pointing to None is different from a name which does not exist 'qq'
'pp' in globals()
'qq' in globals()
try:
    print(qq)
except NameError as e:
    print(e) #e.args type(e)

#
# Usage inside if conditional: None is falsy (taken as FALSE), use 'is None' 
if(pp):
    print("Variable is None")
else:
    print("Variable is Not None")

#
```

## Indexing

- Indexing starts at 1 in R. Whereas, it starts at 0 in Python

## Pipe

- Python uses dot `.` as pipe operator
- R uses `|>` as pipe operator along with underscore `_` as placeholder
  - To minimize the issues, in this book, underscore will be used (and dot will be avoided) in names of objects or variables names

## Assignment Operator

- Python uses 'equal to' ` = ` as assignment operator 
- R uses 'left arrow' ` <- ` as assignment operator
  - In R, while the ` = ` can be used for assignment, its usage for assignment is highly discouraged because it may behave differently under certain subtle conditions which are difficult to debug
  - Convention is to use ` = `  only during function calls for arguments association (syntactic token)

## Copy Objects or Variables

- R [Advanced R, Hadley](https://adv-r.hadley.nz/names-values.html "https://adv-r.hadley.nz/names-values.html")

- Python Similar but different
  
```{r 'R-Copy', decorate=TRUE}
if(FALSE) library("lobstr")
# Create an object containing a value and bind that object to name 'aa'
aa <- 10 
# Copy an object i.e. Create another binding to an already existing value
bb <- aa
# Note that both names are pointing to same memory address
obj_addr(aa)
obj_addr(bb)
stopifnot(identical(obj_addr(aa), obj_addr(bb)))

# Print the copy
print(bb)
# Modify the copy
bb <- 5
# Print the copy and the original (original is unmodified)
print(bb)
print(aa)
# Now the modified name points to a different memory address than earlier
obj_addr(bb)
# Original is still pointing to the same address containing original object
obj_addr(aa)
```

```{python 'Y-Copy', decorate=TRUE}
# Create a variable containing a value and bind that variable to name 'pp'
pp = 10 
# Copy a variable i.e. Create another binding to an already existing value
qq = pp
# Note that both names are pointing to same memory address
id(pp)
id(qq)
assert(id(pp) == id(qq))

# Print the copy
print(qq)
# Modify the copy
qq = 5
# Print the copy and the original (original is unmodified)
print(qq)
print(pp)
# Now the modified name points to a different memory address than earlier
id(qq)
# Original is still pointing to the same address containing original variable
id(pp)

```

## Reticulate Type conversion

- Refer Table \@ref(tab:P01T01) for conversion rules
  - Python variables are available to R as the elements of `py` object 
  - R objects are available to Python as the elements of `r` class

Table: (\#tab:P01T01) (P01T01) Reticulate Type conversion

| R | Python |
| :--- | :--- | 
| Single element vector | Scalar |
| Multi element vector | List |
| List of multiple types | Tuple |
| Named List | Dictionary |
| Matrix or Array | NumPy ndarray |
| Data Frame | Pandas Data Frame |


## Setup Python with Reticulate in R

```{r 'R-Load-Python', decorate=TRUE, eval=FALSE}
if(FALSE) {
  # In R package::reticulate handles Python
  library("reticulate")
  
  # Provide the path to the specific Python binary.
  use_python("C:\\Softwares\\Python\\Python310\\python.exe", required = TRUE)
  
  # PATH: C:\Softwares\Python\Python310\python.exe
  Sys.which("python")
}
```

## Working Directory

```{r 'R-Directory', decorate=TRUE}
getwd()             # Working Directory
```

```{python 'Y-Directory', decorate=TRUE}
os.getcwd()         # Working Directory

```

## Verify R & Python Installation

```{r 'R-Version', decorate=TRUE}
strsplit(R.version.string, " ")[[1]][3]           # R Version
packageVersion("knitr")                           # Package Version: knitr
```

- Python (in PowerShell Terminal)
  - If the cmd can locate python correctly but PowerShell cannot, then execute following  
    - `$env:path="$env:Path;C:\Softwares\Python\Python310"`
  - If the python command leads to Microsoft Store
    - Windows | Start | Manage app execution aliases | Disable two for 'python.exe' and 'python3.exe'
    - It hides access to the actual exe PATH because it comes before the actual 
  - The executable gives different path depending upon from where it is being executed
    - Python: `C:\\Softwares\\Python\\Python310`
    - Local RMarkdown: `C:\\Program Files\\RStudio\\bin`
    - Knit: `C:\\PROGRA~1\\R\\R-42~1.1\\bin\\x64`

```{bash 'Z-Python-Version'}
# Check, in BASH, if the python has been installed 
python --version

# PATH
where.exe python

# Get the pip Version
bool_as_str=true;
if [[ "$bool_as_str" == "true" ]]; then pip --version; fi

```

```{python 'Y-Executable'}
# Sys Executable Path
os.path.dirname(sys.executable)

print(sys.version.split()[0])           # Python Version
print(pd.__version__)                   # Module Version: pandas

```

## Install & Update of R Packages and Python Modules

```{r 'R-Package-Install', decorate=TRUE, eval=FALSE}
if(FALSE) {
  # Install R Packages  
  install.packages("knitr", dependencies = TRUE)
  
  # Find whether a package is installed or not (slow)
  installed.packages() |> rownames() |> is.element(el = "knitr", set = _)
}
```

- R: Use HMI
  - RStudio | Tools | Check for Package Updates | Select | Install Updates
    - Packages which are 'loaded via a namespace' by RStudio on start-up (e.g. `knitr`), can be updated in R
  - R | Packages | Update Packages | Select | OK
    - Packages which are bundled with R (e.g. `MASS`), can be updated by starting R in Administrator Mode.

- Python: Use `pip` (in a PowerShell Terminal)
  - Install modules : `pip install pandas`
  - Upgrade `pip` itself : `python -m pip install --upgrade pip`
  - Upgrade other modules :
    - Get a List of Outdated Modules in Freeze Format, split to get only Module Names, upgrade
    - `pip list --outdated --format=freeze | %{$_.split("==")[0]} #| %{pip install --upgrade $_}`
    - Note: Upgrade step of the pipe has been commented out to prevent accidents.
  - Find whether a module is installed or not : `pip list | grep pandas`

## Attach an R Package & Load a Python Module

```{r 'R-Package-Attach', decorate=TRUE, eval=FALSE}
if(FALSE) {
  # Attach an R Package
  library(knitr)
  
  # Attach multiple R Packages
  c("reticulate", "knitr") |> lapply(library, character.only=TRUE)
  
  # List all attached Packages
  print(.packages())
}
```

```{python 'Y-Module-Load', decorate=TRUE}
if(False):
    # Load a Python Module
    import sys
    import pandas as pd
    import numpy as np
    
    # Load multiple Python Modules
    import sys, pandas as pd, numpy as np
    
    # List all imported Modules
    # Works except the implicit import: 'from x import y' (avoid this)
    # Use k to list alias or use v.__name__ to list actual module names
    allmodules = [v.__name__ for k, v in globals().items() 
        if type(v) is types.ModuleType and not k.startswith('__')]
    
    len(allmodules)
    
    ", ".join(allmodules)

```

## R sessionInfo()

- R `sessionInfo()` provides a quick look at OS, locale, and the packages loaded
- Python does not have similar in-built method. However,
  - `pip list` can be used along with `globals()`
  - Module `session_info` attempts to provide similar output

```{r 'R-sessionInfo', decorate=TRUE, eval=FALSE}
if(FALSE) {
  # List of all attached packages and OS related information
  sessionInfo()
  
  # Base Packages
  aa <- c("stats","graphics","grDevices","datasets","utils","methods","base")
  stopifnot(identical(setdiff(sessionInfo()$basePkgs, aa), character(0)))
  
  # other attached packages
  names(sessionInfo()$otherPkgs)
  
  # loaded via a namespace (and not attached)
  names(sessionInfo()$loadedOnly)
}
```




## Verify

```{r 'R-List-Objects', decorate=TRUE}
ls()
```
