# Data Structures {#structures-p02}

```{r 'setup', include=FALSE}
################################################################################
source("_common.R")
if(FALSE) c("lobstr") |> 
  lapply(library, character.only=TRUE)
```

## R Atomic Vectors or Data Types

- R has 6 basic data types (`logical`, `integer`, `double`, `character`, `complex`, and `raw`). These data types can be combined to form Data Structures (`vector`, `list`, `matrix`, `dataframe`, `factor`). 
  - Vectors are the simplest type of data structure in R. A `vector` is a sequence of data elements of the same basic type. 
  - Members of a `vector` are called 'components'.
  - Atomic vectors are homogeneous i.e. each component has the same datatype. 
  - A vector type can be checked with the `typeof()` function.
  - `list` is a `vector` but not an 'atomic vector'.

- Create a vector or a list by `c()`
  - In R, a literal character or number is just a vector of length 1. 
  - So, `c()` 'combines' them together in a series of 1-length vectors. It neither 'creates' nor 'concatenates' the vectors. It combines lists into a list and vectors into a vector.
  - All attributes (e.g. `dim`) except `names` are removed.
  - All arguments are coerced to a common type 
  - The output type is determined from the highest type of the components in the hierarchy `NULL` < `raw` < `logical` < `integer` < `double` < `complex` < `character` < `list` < `expression`.
  
```{r 'R-DataTypes', decorate=TRUE}
# Integer: To declare as integer "L" (not "l") is added as Suffix
str(c(1L, 2L, NA, 4L, 5L))

# Double (& Default)
str(c(1, 2, NA, 4, 5))

# Character
str(c('a', 'b', NA, 'd', 'e'))

# Logical
str(c(TRUE, FALSE, NA, FALSE, TRUE))
```

## Examination of R Data Types

```{r 'R-About', decorate=TRUE}
# To know about an Object Named Vector (pi, letters are predefined)
aa <- setNames(c(1, 2, NA, pi, 4), nm = letters[1:5])

typeof(aa)              # Type
class(aa)               # Class
str(aa)                 # Structure
length(aa)              # Length
dim(aa)                 # Dimensions
is(aa)                  # Inheritance
names(attributes(aa))   # Attributes
names(aa)               # Names
```

## R Matrices and Arrays

- `Matrices` and `arrays` are simply vectors with the attribute `dim` attached to the vector. 
  - The data elements must be of the same basic type.
  - A `matrix` is a two-dimensional rectangular data set. 
  - `Arrays` are multi-dimensional Data structures. Data is stored in the form of matrices, row, and as well as in columns. 

```{r 'R-Matrix', decorate=TRUE}
# Create Matrix
aa <- matrix(1:6, nrow=3, ncol=2, byrow=TRUE, dimnames=list(NULL, c("x", "y")))
bb <- matrix(1:6, nrow=3, ncol=2, byrow=FALSE, dimnames=list(NULL, c("x", "y")))

print(aa)
print(bb)

# About
str(aa)
dim(aa)
length(aa)

# Matrices have 'dimnames' attribute instead of usual 'names'
names(attributes(aa))
names(aa)
dimnames(aa) 
```

## R Lists

- Lists have 'elements', each of which can contain any type of R object, i.e. the elements of a list do not have to be of the same type.
  - `list` is a `vector` but not an 'atomic vector'.

```{r 'R-List', decorate=TRUE}
# Create List (letters is predefined vector of 26 elements)
aa <- list(x = 4:6, y = letters[4:8])
print(aa)

# About
str(aa)
length(aa)              # Number of elements in the List
lengths(aa)             # Number of components of each element of the List
stopifnot(identical(lengths(aa), sapply(aa, length)))

names(attributes(aa))   # Attributes
names(aa)               # Names of elements
```

## R DataFrames

- Data frame is a list of `vectors`, `factors`, and/or `matrices` all having the same length (number of rows in the case of matrices).
  - A `data frame` can contain a `list` that is the same length as the other components. 

```{r 'R-DataFrame', decorate=TRUE}
# Create DataFrame (letters is predefined vector of 26 elements)
aa <- data.frame(x = 4:6, y = letters[4:6])
print(aa)

# About
str(aa)
dim(aa)                 #Dimensions Row x Column
stopifnot(all(identical(nrow(aa), dim(aa)[1]),
              identical(ncol(aa), dim(aa)[2])))

names(attributes(aa))   # Attributes
names(aa)               # Names of column headers

is.list(aa)
is.vector(aa)
is.atomic(aa)
```

## R Factors

- `Factors` are used to describe items that can have a finite number of values (gender, social class, etc.). A factor has a `levels` attribute and class `factor`. 
  - A factor may be purely nominal or may have ordered categories. 

```{r 'R-Factor', decorate=TRUE}
# Create Factors Unordered
aa <- factor(c("female", "male", "male", "female", "male"), ordered = FALSE)
# Create Factors Ordered
bb <- factor(c("female", "male", "male", "female", "male"), ordered = TRUE)

print(aa)
print(bb)

# About
str(aa)
str(bb)

nlevels(aa)             # Count of Levels
levels(aa)              # Vector of Levels
names(attributes(aa))   # Attributes
```

## Membership Functions

- `anyNA()` is `TRUE` if there is an `NA` present, `FALSE` otherwise
- `is.atomic()` is `TRUE` for all atomic vectors, factors, matrices but is `FALSE` for lists and dataframes
- `is.vector()` is `TRUE` for all atomic vectors, lists but is `FALSE` for factors, matrices, DATE & POSIXct
  - It returns `FALSE` if the vector has attributes (except `names`) ex: DATE, POSIXct, DataFrames (even though a Dataframe is a list and a list is a vector)
- `is.numeric()` is `TRUE` for both `integer` and `double`
- `is.integer()`, `is.double()`, `is.character()`, `is.logical()` are `TRUE` for their respective datatypes only
- `is.factor()`, `is.ordered()` are membership functions for factors with or without ordering

```{r 'R-is', decorate=TRUE}
# Create Objects
aa_num <- setNames(c(1, 2, NA, pi, 4), nm = letters[1:5])
bb_mat <- matrix(1:6, nrow=3, ncol=2, byrow=TRUE)
dd_dft <- data.frame(x = 4:6, y = letters[4:6])
ee_lst <- list(x = 4:6, y = letters[4:8])
ff_fct <- factor(c("female", "male", "male", "female", "male"), ordered = FALSE)

# List of Objects
gg <- list(Vector = aa_num, Matrix = bb_mat, DataFrame = dd_dft, 
            List = ee_lst, Factor = ff_fct)

# Apply a membership function on all of the objects inside the list
names(which(sapply(gg, is.atomic)))
names(which(sapply(gg, is.vector)))
names(which(sapply(gg, is.matrix)))
names(which(sapply(gg, is.list)))
names(which(sapply(gg, is.data.frame)))
names(which(sapply(gg, is.factor)))
```


