# Data Structures {#structures-p02}

```{r 'setup', include=FALSE}
source('_common.R')
```

## R Data Types

- R has 6 basic data types (`logical`, `integer`, `double`, `character`, `complex`, and `raw`). These data types can be combined to form Data Structures (`vector`, `list`, `matrix`, `dataframe`, `factor`). 
  - Vectors are the simplest type of data structure in R. A `vector` is a sequence of data elements of the same basic type. 
  - Members of a `vector` are called 'elements'.
  - Atomic vectors are homogeneous i.e. each component has the same datatype. 
  - A vector type can be checked with the `typeof()` function.
  - `list` is a `vector` but not an 'atomic vector'.

- Create a vector or a list by `c()`
  - In R, a literal character or number is just a vector of length 1. 
  - So, `c()` 'combines' them together in a series of 1-length vectors. It neither 'creates' nor 'concatenates' the vectors. It combines lists into a list and vectors into a vector.
  - All attributes (e.g. `dim`) except `names` are removed.
  - All arguments are coerced to a common type 
  - The output type is determined from the highest type of the components in the hierarchy `NULL` < `raw` < `logical` < `integer` < `double` < `complex` < `character` < `list` < `expression`.
  
```{r 'R-DataTypes', decorate=TRUE}
# Integer: To declare as integer 'L' (not 'l') is added as Suffix
str(c(1L, 2L, NA, 4L, 5L))

# Double (& Default)
str(c(1, 2, NA, 4, 5))

# Character
str(c('a', 'b', NA, 'd', 'e'))

# Logical
str(c(TRUE, FALSE, NA, FALSE, TRUE))
```

- Examination of R Data Types

```{r 'R-About', decorate=TRUE}
# To know about an Object Named Vector (pi, letters are predefined)
aa <- setNames(c(1, 2, NA, pi, 4), nm = letters[1:5])

typeof(aa)              # Type
class(aa)               # Class
str(aa)                 # Structure
length(aa)              # Length
dim(aa)                 # Dimensions
is(aa)[1:6]             # Inheritance
names(attributes(aa))   # Attributes
names(aa)               # Names
```

## R Matrices

- `Matrices` and `arrays` are vectors with the attribute `dim` attached to them 
  - The data elements must be of the same basic type.
  - A `matrix` is a two-dimensional rectangular data set. 
  - 'Arrays' are multi-dimensional Data structures. Data is stored in the form of matrices, row, and as well as in columns. 

```{r 'R-Matrix', decorate=TRUE}
# Create Matrix
aa <- matrix(1:6, nrow=3, ncol=2, byrow=TRUE, dimnames=list(NULL, c('x', 'y')))
bb <- matrix(1:6, nrow=3, ncol=2, byrow=FALSE, dimnames=list(NULL, c('x', 'y')))

print(aa)
print(bb)

# About
str(aa)
dim(aa)
length(aa)

# Matrices have 'dimnames' attribute instead of usual 'names'
names(attributes(aa))
names(aa)
dimnames(aa) 
```

## R DataFrames

- Data frame is a list of `vectors`, `factors`, and/or `matrices` all having the same length (number of rows in the case of matrices).
  - A `data frame` can contain a `list` that is the same length as the other components. 

```{r 'R-DataFrame', decorate=TRUE}
# Create DataFrame (letters is predefined vector of 26 elements)
aa <- data.frame(x = 4:6, y = letters[4:6])
print(aa)

# About
str(aa)
dim(aa)                 #Dimensions Row x Column
stopifnot(all(identical(nrow(aa), dim(aa)[1]),
              identical(ncol(aa), dim(aa)[2])))

names(attributes(aa))   # Attributes
names(aa)               # Names of column headers

is.list(aa)
is.vector(aa)
is.atomic(aa)
```

## R Factors

- `Factors` are used to describe items that can have a finite number of values (gender, social class, etc.). A factor has a `levels` attribute and class `factor`. 
  - A factor may be purely nominal or may have ordered categories. 

```{r 'R-Factor', decorate=TRUE}
# Create Factors Unordered
aa <- factor(c('female', 'male', 'male', 'female', 'male'), ordered = FALSE)
# Create Factors Ordered
bb <- factor(c('female', 'male', 'male', 'female', 'male'), ordered = TRUE)

print(aa)
print(bb)

# About
str(aa)
str(bb)

nlevels(aa)             # Count of Levels
levels(aa)              # Vector of Levels
names(attributes(aa))   # Attributes
```

## R Membership

- `anyNA()` is `TRUE` if there is an `NA` present, `FALSE` otherwise
- `is.atomic()` is `TRUE` for all atomic vectors, factors, matrices but is `FALSE` for lists and dataframes
- `is.vector()` is `TRUE` for all atomic vectors, lists but is `FALSE` for factors, matrices, DATE & POSIXct
  - It returns `FALSE` if the vector has attributes (except `names`) ex: DATE, POSIXct, DataFrames (even though a Dataframe is a list and a list is a vector)
- `is.numeric()` is `TRUE` for both `integer` and `double`
- `is.integer()`, `is.double()`, `is.character()`, `is.logical()` are `TRUE` for their respective datatypes only
- `is.factor()`, `is.ordered()` are membership functions for factors with and without ordering respectively

```{r 'R-is', decorate=TRUE}
# Create Objects
aa_num <- setNames(c(1, 2, NA, pi, 4), nm = letters[1:5])
bb_mat <- matrix(1:6, nrow=3, ncol=2, byrow=TRUE)
dd_dft <- data.frame(x = 4:6, y = letters[4:6])
ee_lst <- list(x = 4:6, y = letters[4:8])
ff_fct <- factor(c('female', 'male', 'male', 'female', 'male'), ordered = FALSE)

# List of Objects
gg <- list(Vector = aa_num, Matrix = bb_mat, DataFrame = dd_dft, 
            List = ee_lst, Factor = ff_fct)

# Apply a membership function on all of the objects inside the list
names(which(sapply(gg, is.atomic)))
names(which(sapply(gg, is.vector)))
names(which(sapply(gg, is.matrix)))
names(which(sapply(gg, is.list)))
names(which(sapply(gg, is.data.frame)))
names(which(sapply(gg, is.factor)))
```

## Python Types

- [(doc) Built-in Types](https://docs.python.org/3/library/stdtypes.html#)

- General
  - The principal built-in types are numerics, sequences, mappings, classes, instances and exceptions.
  - Some collection classes are `mutable`. The methods that add, subtract, or rearrange their members in place, and do not return a specific item, never return the collection instance itself but `None`.
  - Practically all objects can be compared for equality, tested for truth value, and converted to a string.

- Truth Value Testing
  - constants defined to be false: `None` and `False`.
  - zero of any numeric type: `0, 0.0, 0j, Decimal(0), Fraction(0, 1)`
  - empty sequences and collections: `'', (), [], {}, set(), range(0)`

- Boolean Operations in ascending order: `and`, `or`, `not`
- There are eight comparison operations in Python.

- Numeric Types: `int`, `float`, `complex`
  - Booleans are a subtype of integers
  - Python Integers have unlimited precision. Whereas, R integers are limited to $(2^{31} - 1 = 2147483647)$

- Iterator Types
  - Sequences always support the iteration methods.

- Sequence Types: `list`, `tuple`, `range`
  - Negative value of index is relative to the end of the sequence in Python. Whereas, it acts to exclude those indices
  - Concatenating immutable sequences always results in a new object. 
  - The `range` type represents an immutable sequence of numbers and is commonly used for looping a specific number of times in `for` loops.

- Text Sequence Type: `str`
  - Strings are immutable sequences of Unicode code points.

- Set Types: `set` (mutable), `frozenset` (immutable)

- Mapping Types: `dict`

## Similarities

```{r 'R-Outcome-Same', decorate=TRUE}
# R results are same as Python
stopifnot(0  == 1 %/% 2 )
stopifnot(-1 == (-1) %/% 2 )
stopifnot(-1 == 1 %/% (-2) )
stopifnot(0  == (-1) %/% (-2) )

# 1^y and y^0 are 1, ALWAYS in both R and Python
stopifnot(all(sapply(list(0**0, NaN**0, NA**0, Inf**0, 1**NA, 1**NaN), 
                     identical, 1)))

stopifnot(is.nan(NaN * 0))
stopifnot(is.na(NA * 0) & !is.nan(NA * 0))        #NA

stopifnot(is.nan(Inf * 0))
```

```{python 'Y-Outcome-Same', decorate=TRUE}
# Python results are same as R
assert(0  == 1//2 )
assert(-1 == (-1) // 2 )
assert(-1 == 1 // (-2) )
assert(0  == (-1) // (-2) )

# 1^y and y^0 are 1, ALWAYS in both R and Python
assert(1 == 0**0 == np.nan**0 == 1**np.nan)

assert(np.isnan(np.nan * 0))
assert(np.isnan(math.inf * 0))

```

## Differences

```{r 'R-Outcome-Different', decorate=TRUE}
# R allows division by zero whereas Python throws ZeroDivisionError
stopifnot(is.na(0L %/% 0L) & !is.nan(0L %/% 0L))  #NA
stopifnot(is.nan(0 %/% 0 ))                       #NaN
stopifnot(is.infinite(1 %/% 0))                   #Inf
```

```{r 'R-Sequence', decorate=TRUE}
# R index starts from 1, whereas Python index starts from 0
# R range includes the maximum value, whereas Python excludes it
seq.int(length.out = 5)
seq.int(from = 0, length.out = 5)
seq.int(to = 10, length.out = 5)
seq.int(by = -2, length.out = 5)

# Colon ':' acts as sequence in R, Python uses colon for dictionary key:value
# Colon should be avoided in R if the range limits may change
1:5
1:0
```

```{python 'Y-Sequence', decorate=TRUE}
# Python index starts from 0, whereas R index starts from 1
# Python range excludes the maximum value, whereas R includes it
list(range(5))
list(range(1, 6))
list(range(10, 1, -2))

```

## Python Collections

- Python Collections: List, Tuple, Set, Dictionary
  - `tuple`: Literal `()`, Ordered, Immutable, Allows Duplicates, [Refer](https://docs.python.org/3/library/stdtypes.html#tuples)
  - `list` : Literal `[]`, Ordered, Mutable, Allows Duplicates, [Refer](https://docs.python.org/3/library/stdtypes.html#lists)
  - `dict` : Literal `{}`, Ordered, Mutable, No Duplicates, [Refer](https://docs.python.org/3/library/stdtypes.html#mapping-types-dict)
  - `set`  : Literal `{}`, Unordered, Mutable, No Duplicates, [Refer](https://docs.python.org/3/library/stdtypes.html#set-types-set-frozenset)
    - Effectively, 'sets' are 'dictionaries' without keys
    - Indexing and slicing does not work on `set` because it is unordered
    - Immutable objects include `numbers`, `strings` and `tuples`. [Refer](https://docs.python.org/3/glossary.html). 
    - `set` is converted to `R environment`. However, it is tricky to handle, so it is being put on hold for now. Note that python set operations like 'set difference' can be applied using scope resolution i.e. `py$set_x$difference(set_y)`


```{python 'Y-Lists', decorate=TRUE}
pp = 11                                           # int
print(f'{type(pp) = }   |     ...     | {pp = }')

pp = 11,                                          # Implicit Tuple
print(f'{type(pp) = } | {len(pp) = } | {pp = }')

pp = (11, )                                       # Length 1 Tuple needs comma
print(f'{type(pp) = } | {len(pp) = } | {pp = }')

pp = (11, 22, 33)                                 # Tuple
print(f'{type(pp) = } | {len(pp) = } | {pp = }')

pp = [11, 22, 33]                                 # List
print(f'{type(pp) = }  | {len(pp) = } | {pp = }')

pp = {11, 22, 33}                                 # Set (unordered)
print(f'{type(pp) = }   | {len(pp) = } | {pp = }')

pp = {'a': 11, 'b': 22, 'c': 33}                  # Dictionary
print(f'{type(pp) = }  | {len(pp) = } | {pp = }')

```

## NumPy Array

```{python 'Y-NumPy-Array', decorate=TRUE}
pp = np.arange(12).reshape((3,4))
print(type(pp))
print(pp)
assert(np.array_equal(pp[0, :], pp[0, ]))         #Verify same shape & values
print(pp[0, ])                                    #Subset First Row

```

