# R List & Python Dictionary {#list-dict-p03}

```{r 'setup', include=FALSE}
source('_common.R')
```


## Basics

- R `list` : [CRAN Lists](https://cran.r-project.org/doc/manuals/R-intro.html#Lists)
  - An R `list` is an object consisting of an ordered collection of objects known as its 'components'. 
    - Each component can be of any type of R object and of any length. It can even include `list`, `environment` etc.
  - Components are always numbered. Components may also be named.
  - The vector of `names` is in fact simply an attribute of the list.
  - $[[ \ldots ]]$ vs. $[ \ldots ]$
    - `[[ ]]` is the operator used to select a single element, whereas `[ ]` is a general subscripting operator. 
    - Thus `[[ ]]` provides the first object in the list, and if it is a named list the `name` is not included. 
    - However, `[ ]` provides a sublist of the list consisting of the first entry only. If it is a named list, the `names` are transferred to the sublist.
  - `list` is a `vector` but not an 'atomic vector'

- Python `dict` : [Python Dictionary](https://docs.python.org/3/library/stdtypes.html#mapping-types-dict)
  - `dict` : Literal `{}`, Ordered, Mutable, No Duplicates
  - Dictionaries compare equal if and only if they have the same (key, value) pairs (regardless of ordering).
  - Essentially it means 'insertion ordered' which is different from ordered like `OrderedDict`
  - Equality comparisons with `OrderedDict` are order sensitive but this is not the case for `dict`


- Create & Print 

```{r 'R-List-Create', decorate=TRUE}
aa = list()                                       #Empty List
print(aa)

aa <- list(x = c(11L, 12L), y = c('a', 'b', 'c')) #Simple List
print(aa)

str(aa)                                           #Structure
```

```{python 'Y-Dict-Create', decorate=TRUE}
pp = {}                                           #Empty Dictionary
print(f'{type(pp) = } | {len(pp) = } | {pp = }')


pp = {'x': [11, 12], 'y': ('a', 'b', 'c')}        #Simple Dictionary
print(f'{len(pp)=}, {pp = }')

```

- Order sensitive comparison is done in R for `list`. However, Python does not consider position during `dict` comparison

```{r 'R-List-Compare', decorate=TRUE}
# Create list by separately supplying component names
aa <- list(x = c(11L, 12L), y = c('a', 'b', 'c'))
bb <- setNames(list(c(11L, 12L), c('a', 'b', 'c')), nm = c('x', 'y'))
stopifnot(identical(aa, bb))

dd <- list(y = c('a', 'b', 'c'), x = c(11L, 12L))
stopifnot(!identical(aa, dd))           #FALSE: R considers position
#To ignore position (like Python)
stopifnot(identical(aa[order(names(aa))], dd[order(names(dd))]))

```

```{python 'Y-Dict-Compare', decorate=TRUE}
# Create dict by separately supplying component names
pp = {'x': [11, 12], 'y': ('a', 'b', 'c')}
qq = dict(zip(['x', 'y'], [[11, 12], ('a', 'b', 'c')]))

ss = {'y': ('a', 'b', 'c'), 'x': [11, 12]}
assert(pp == qq == ss)                  #True: Python does not consider position
assert(list(pp.items()) != list(ss.items()))      #To consider position (like R)

```

- Count Components

```{r 'R-List-Count', decorate=TRUE}
aa <- list(x = c(11L, 12L), y = c('a', 'b', 'c')) #Simple List
length(aa)                                        #Count components

lengths(aa)                                       #Count elements of components
```

```{python 'Y-Dict-Count', decorate=TRUE}
pp = {'x': [11, 12], 'y': ('a', 'b', 'c')}        #Simple Dictionary
len(pp)                                           #Count key: value pairs


{k:len(v) for k, v in pp.items()}                 #Count length of each pair

```

- Names and Types

```{r 'R-List-Types', decorate=TRUE}
aa <- list(x = c(11L, 12L), y = c('a', 'b', 'c')) #Simple List
names(attributes(aa))                             #Attributes

bb <- names(aa)                                   #Names of components
bb

# Find Names
stopifnot(all( 'x' %in% bb,  any('x' == bb),  is.element('x', set = bb), 
              !'z' %in% bb, !any('z' == bb), !is.element('z', set = bb))) 

# Note: %in% operator is NOT commutative i.e. a %in% b != b %in% a
stopifnot(! identical(1 %in% c(1, 2), c(1, 2) %in% 1))

sapply(aa, typeof)                                #Type of components
```

```{python 'Y-Dict-Types', decorate=TRUE}
pp = {'x': [11, 12], 'y': ('a', 'b', 'c')}        #Simple Dictionary
list(pp.keys())                                   #List of keys

assert('x' in pp.keys() and 'z' not in pp.keys()) #Find Keys

list(pp.values())                                 #List of Values


list(pp.items())                                  #List of items


print('\n'.join([k + ' : ' + str(type(pp[k]).__name__) 
                for k, v in pp.items()]))         #Type of items

```

- Slice or Subset or Select: Indexing starts from 1 in R and from 0 in Python.

```{r 'R-List-Select', decorate=TRUE}
aa <- list(x = c(11L, 12L), y = c('a', 'b', 'c')) #Simple List

bb <- aa[[1]]                                     #[[ ]] First Object: index 1
stopifnot(all(sapply(list(aa$x, aa[['x']]), identical, bb)))
print(bb)
str(bb)

dd <- aa[1]                                       #[ ] Sublist with First Entry
print(dd)
str(dd)
```

```{python 'Y-Dict-Select', decorate=TRUE}
pp = {'x': [11, 12], 'y': ('a', 'b', 'c')}        #Simple Dictionary

qq = list(pp.items())[0]                          #First Object: index 0
print(f'{type(qq) = } | {len(qq) = } | {qq = }')  #Converted to Tuple


qq = pp['x']                                    
print(f'{type(qq) = } | {len(qq) = } | {qq = }')  #Keep original as list

```

- Type checking

```{r 'R-List-Class', decorate=TRUE}
aa <- list(x = c(11L, 12L), y = c('a', 'b', 'c')) #Simple List
stopifnot(is.list(aa))

stopifnot(is.vector(aa))                          #TRUE: list is vector

stopifnot(!is.atomic(aa))                         #FALSE: list is not atomic
```

```{python 'Y-Dict-Class', decorate=TRUE}
pp = {'x': [11, 12], 'y': ('a', 'b', 'c')}        #Simple Dictionary

assert(pp.__class__ is dict)                      #same as type() (Avoid)

assert(type(pp) is dict)                          #type() (Strict)

assert(isinstance(pp, dict))                      #isinstance() allows subtypes

assert(isinstance(False, int))                    #True: bool is subtype of int
assert(type(False) is not int)

```

- Merge or Combine in R allows duplicate components but Python overwrites them (Last seen wins) 

```{r 'R-List-Merge', decorate=TRUE}
aa <- list('a' = 11L, 'b' = 22L)
bb <- list('c' = 33L, 'b' = 222L)

dd <- c(aa, bb)                         #R keeps duplicates (Avoid)
str(dd)

ee <- modifyList(aa, val = bb)          #To overwrite duplicates (like Python)
str(ee)

#Overwrite duplicates: This may result in position change of key
str(c(aa, bb)[!duplicated(c(names(aa), names(bb)), fromLast = TRUE)])

```

```{python 'Y-Dict-Merge', decorate=TRUE}
pp = {'a': 11, 'b': 22}
qq = {'c': 33, 'b': 222}

ss = copy.deepcopy(pp)
ss.update(qq)                           #Update in-place

tt = {**pp, **qq}                       #Merge using **

uu = pp | tt                            #Merge using |

vv = copy.deepcopy(pp)
vv |= qq                                #Update in-place using |=
assert(ss == tt == uu == vv)

print(f'{len(ss)=}, {ss = }')

```

- Add or Update Components

```{r 'R-List-Update', decorate=TRUE}
aa <- list('a' = 11L, 'b' = 22L)

aa$b <- 222L
aa$b

aa[['b']] <- 22L

aa$c <- 33L
str(aa)
```

```{python 'Y-Dict-Update', decorate=TRUE}
pp = {'a': 11, 'b': 22}

pp['b'] = 200                                     #subscript notation
pp['c'] = 33
assert(pp['b'] == 200)

pp.update({'b': 300})                             #Slow
assert(pp['b'] == 300)

pp.update([('b', 22)])                            #Slow
assert(pp['b'] == 22)

print(f'{len(pp)=}, {pp = }')

```

- Delete components

```{r 'R-List-Delete', decorate=TRUE}
aa <- list('a' = 11L, 'b' = 22L, 'c' = 33L)

bb <- within(aa, rm('b', 'c'))                    #Error if key not found

bb[c('b', 'c', 'xyz')] <- NULL                    #No error if key not found

dd <- bb[setdiff(names(aa), c('b', 'c', 'xyz'))]  #No error if key not found
stopifnot(identical(bb, dd))
str(bb)
```

```{python 'Y-Dict-Delete', decorate=TRUE}
pp = {'a': 11, 'b': 22, 'c': 33}
qq = copy.deepcopy(pp)
ss = copy.deepcopy(pp)
tt = copy.deepcopy(pp)

uu = {k: v for k, v in pp.items() if k not in ('b', 'c', 'xyz')} #Slow

for k in ('b', 'c'): del qq[k]          #Return None, Error if key not found

for k in ('b', 'c'): ss.pop(k)          #Return Value, Error if key not found
for k in ('xyz'): ss.pop(k, None)       #Using None, No error if key not found

[tt.pop(k) for k in ('b', 'c')]
[tt.pop(k, None) for k in ('xyz')]


assert(qq == ss == tt == uu)
print(f'{len(qq)=}, {qq = }')

```

## Nested

```{r 'R-List-Nest', decorate=TRUE}
aa <- list('a' = 11L, 'b' = 22L)
bb <- list('c' = 33L, 'd' = 44L, 'e' = aa)        #Nested List
str(bb)
```

```{python 'Y-Dict-Nest', decorate=TRUE}
pp = {'a': 11, 'b': 22}
qq = {'c': 33, 'd': 44, 'e': pp}                  #Nested Dictionary
print(f'{len(qq)=}, {qq = }')

```

## With Multiple Types

```{r 'R-List-Multi', decorate=TRUE}
aa_lst <- list(11L, 22L)                #Unnamed List   | Tuple
bb_lst <- list(a = 11L, b = 22L)        #Named List     | Dict
dd_vec <- c(11L, 22L)                   #Unnamed Vector | List

ee <- list(a = 11L, b = 'x', c = FALSE, d = NULL, 
           e = aa_lst, f = dd_vec, g = bb_lst)
lst_r <- ee                             #R list to transform into Python dict
#str(ee)
```

```{python 'Y-Dict-Multi', decorate=TRUE}
# Exclude Python sets because these are converted to R Environment
pp_tup = (11, 22)             #Tuple: Ordered, Immutable | Unnamed List
qq_lst = [11, 22]             #List: Ordered, Mutable    | Unnamed Vector
ss_dic = {'a': 11, 'b': 22}   #Dict: Ordered, Mutable    | Named List

tt = {'a': 11, 'b': 'x', 'c': False, 'd': None, 
      'e': pp_tup, 'f': qq_lst, 'g': ss_dic}
dic_y = copy.deepcopy(tt)     #Python dict to transform into R list

#print('\n'.join([k + ' : ' + str(type(tt[k]).__name__) + ' : ' + str(v) 
#                for k, v in tt.items()]))

```

```{r 'R-List-Transform', decorate=TRUE}
aa <- py$dic_y                          #Convert Python dict to R List
stopifnot(identical(lst_r, aa))         #TRUE
str(aa)
```

```{python 'Y-Dict-Transform', decorate=TRUE}
pp = r.lst_r                            #Convert R List to Python dict
pp['e'] = tuple(pp['e'])                #List to Tuple
assert(dic_y == pp)                     #True

print('\n'.join([k + ' : ' + str(type(pp[k]).__name__) + ' : ' + str(v) 
                for k, v in pp.items()]))

```

