# Data Frame {#df-p04}

```{r 'setup', include=FALSE}
source("_common.R")
```

R provides `data.frame` for tabular data structure. `tibble` & `data.table` are packages which extends its capabilities. Python module `pandas` provide `DataFrame` with similar capabilities. 

> Indexing starts from 1 in R and from 0 in Python.  

## Basics

- R `data.frame` is a `list` of variables of the same number of rows. It is a `matrix` like structure whose columns may be of differing types.
- Python Pandas `DataFrame` is a 2-dimensional data structure that can store data of different types in columns. 


- Create R `data.frame()`, `tibble()` and Python `pd.DataFrame()`

```{r 'R-DF-Create', decorate=TRUE}
nn <- 4L                                          #Number of Rows

# R Data Frame: integer, double, character, logical, factor
df_r <- data.frame(
  INT = 1:nn, NUM = seq(1, nn, 1), CHR = letters[1:nn], LGL = (1:nn %% 2) == 0, 
  FCT = factor(rep(c('No', 'Yes'), length.out = nn)))

# Tibble
tbl <- tibble(
  INT = 1:nn, NUM = seq(1, nn, 1), CHR = letters[1:nn], LGL = (1:nn %% 2) == 0, 
  FCT = factor(rep(c('No', 'Yes'), length.out = nn)))

stopifnot(all(identical(df_r, as.data.frame(tbl)), 
              identical(tbl, as_tibble(df_r))))
```

```{python 'Y-DF-Create', decorate=TRUE}
pp = 4                                            #Number of Rows

qq = {'INT': [i+1 for i in range(pp)], 
      'NUM': (float(i+1) for i in range(pp)),
      'CHR': [chr(i) for i in range(ord('a'), ord('a') +pp)],
      'LGL': [i % 2 == 1 for i in range(pp)],
      'FCT': pd.Categorical(['No', 'Yes'] * 2)}

df_y = pd.DataFrame(data = qq)                    #DataFrame from dict

```

```{r 'R-Compare', decorate=TRUE}
aa <- py$df_y |> `attr<-`('pandas.index', NULL)
aa$INT <- as.integer(aa$INT)
stopifnot(identical(df_r, aa))
```

- R: `print()`, `head()`, `tail()`, `slice()`
- Python: `print()`, `pd.head()`, `pd.tail()`, `pd.iloc()`

```{r 'R-DF-Head', decorate=TRUE}
aa <- df_r
if(FALSE) print(df_r)                   #Data Frame prints ALL Rows (Avoid)
if(FALSE) print(tbl, n = 2)             #Tibble can take number of rows to print

stopifnot(identical(slice(aa, 1:2), head(aa, 2)))
head(aa, 2)                                       #Subset by Head


tail(aa, 2)                                       #Subset by Tail
```

```{python 'Y-DF-Head', decorate=TRUE}
pp = df_y.copy()

assert(pp.head(2).equals(pp.iloc[:2]))
pp.head(2)


pp.tail(2)

```

- R: `class()`, `type()`, `dim()`, `names()`, `str()`, `summary()`
- Python: `typeof()`, `pd.shape`, `pd.columns.values`, `pd.describe()`, `pp.dtypes`, `pd.info()`

```{r 'R-DF-About', decorate=TRUE}
aa <- df_r

class(aa)                                         #Class

typeof(aa)                                        #Type

dim(aa)                                           #Dimensions [Row, Column]

names(aa)                                         #Names

str(aa)                                           #Structure

summary(aa)                                       #Summary
```

```{python 'Y-DF-About', decorate=TRUE}
pp = df_y.copy()

print(type(pp))                                   #Explicitly Print type()


pp.shape                                          #Dimensions [Row, Column]


list(pp.columns)                                  #Names


list(pp.describe().index)                         #(Default) Summary of Num only
[x for x in pp.describe(include = 'all').index if x not in pp.describe().index]

pp.describe(include = 'all').loc[['count', 'max', 'unique']]


pp.dtypes                                         #data type of each column


pp.info(memory_usage = False)                     #Structure

#Prevent the collapse of middle columns into (...)
if(False):
    with pd.option_context('display.max_columns', 40):
        print(pp.describe(include = 'all'))

```

- R: `dplyr::select()`, `data.frame()`, `with()`, `[]`, `subset()` (Avoid)
- Python: `pd.[]`, `pd.drop()`

```{r 'R-DF-Select', decorate=TRUE}
aa <- df_r
names(aa)

# dplyr::select() by including or excluding names
bb <- select(aa, 2:3)
dd <- select(aa, NUM, CHR)
ee <- select(aa, -c(INT, LGL, FCT))

ff <- data.frame('NUM' = aa$NUM, 'CHR' = aa$CHR)
# with() can be used to create an environment using data
gg <- with(aa, data.frame(NUM, CHR))
# [] is used for subsetting but note that 1-column subset is vector by default
hh <- aa[ , c('NUM', 'CHR'), drop = FALSE]
ii <- subset(aa, select = c(NUM, CHR), drop = FALSE)      #Avoid

stopifnot(all(sapply(list(dd, ee, ff, gg, hh, ii), identical, bb)))
str(bb)
```

```{python 'Y-DF-Select', decorate=TRUE}
pp = df_y.copy()
list(pp.columns)

qq = pp[['NUM', 'CHR']].copy()                              #Use List of Names
ss = pp.drop(columns = ['INT', 'LGL', 'FCT']).copy()        #Drop Columns
tt = pp.drop(['INT', 'LGL', 'FCT'], axis = 1).copy()        #0 Rows, 1 Columns

assert(qq.equals(ss) and qq.equals(tt))
qq

```

- R: `names()`, `dplyr::rename()`
- Python: `pd.rename()`, `pd.columns.values`

```{r 'R-DF-Rename', decorate=TRUE}
aa <- df_r
names(aa)

names(aa)[c(1, 3)] <- c('A', 'C')                 #By Position
names(aa)[names(aa) == 'NUM'] <- 'B'              #By Name

aa <- rename(aa, D = LGL, E = 5)                  #New = Old (Reverse of Python)
names(aa)
```

```{python 'Y-DF-Rename', decorate=TRUE}
pp = df_y.copy()
list(pp.columns)

pp.rename(columns = {'INT': 'A', pp.columns[1]: 'B'}, inplace = True) #Old: New
pp.rename(columns = dict(zip(pp.columns[[3]], ['D'])),inplace = True) #New, Old
pp.columns.values[[2, 4]] = ['C', 'E']

list(pp.columns)

```

- R: `order()`, `dplyr::arrange()`

```{r 'R-DF-Order', decorate=TRUE}
aa <- df_r

# order() ALWAYS reorders the rownames
aa[order(aa$CHR, decreasing = TRUE), ]


# arrange() reorders character rownames but reinitialises them from 1 if integer
arrange(aa, desc(CHR))
```

```{python 'Y-DF-Order', decorate=TRUE}
pp = df_y.copy()
list(pp.index)


pp.sort_values('CHR', ascending = False, inplace = True)
list(pp.index)

```

## RowNames - Index

- R `rownames` are called `index` in Python
- R `data.frame` has `rownames` but `tibble` heavily discourage their usage. 
- Duplicated row indices are allowed in Python but not in R

```{r 'R-DF-RowNames', decorate=TRUE}
aa <- data.frame(x = 1:2)
row.names(aa)

row.names(aa) <- letters[1:2]                               #rownames

stopifnot(has_rownames(aa))
row.names(aa)
```

```{python 'Y-DF-RowNames', decorate=TRUE}
pp = pd.DataFrame(data = {'x': [1, 2]})
list(pp.index)


pp.set_index([pd.Index(['a', 'a'])], inplace = True)        #Duplicate index
pp.set_index([pd.Index(['a', 'b'])], inplace = True)
list(pp.index) 

```

## Copy

```{r 'R-DF-Copy', decorate=TRUE}
bb <- aa <- data.frame(x = 11:13)                 #R Copy-on-Modify
stopifnot(identical(obj_addr(aa), obj_addr(bb)))
obj_addr(aa)

aa[2, 'x'] <- NA

obj_addr(aa)                  #Points to an address different from earlier
aa$x                          #Modified

obj_addr(bb)                  #No change in address
bb$x
```

```{python 'Y-DF-Copy', decorate=TRUE}
pp = pd.DataFrame({'x': [1, 2, 3]})
qq = pp.copy()                          #Deepcopy by default
ss = copy.deepcopy(pp)                  #Deepcopy

assert(id(pp) != id(qq) != id(ss))      #Deepcopy has different address
id(pp)


pp.at[1, 'x'] = None

assert(qq.equals(ss) and not qq.equals(pp))

id(pp)                                  #Points to the same address as before

pp

```

## Modify [] - .at, .loc

- R `[]` vs. Python `pd.at[]`, `pd.loc[]`, `pd.iat[]`, `pd.iloc[]`

```{r 'R-DF-Modify', decorate=TRUE}
aa <- df_r

aa[1, 2] <- NA
aa[2, 'CHR'] <- NA

aa
```

```{python 'Y-DF-Modify', decorate=TRUE}
pp = df_y.copy()

# .at can only access a single value at a time. Fast
# It tries to maintain the datatype (fails sometimes)
# If column number is used by mistake, it creates a New Column (Avoid)
pp.at[0, 'NUM'] = None

# .loc can select multiple rows and/or columns. Slow
# It does not maintain the datatype and modifies the type silently
pp.loc[1, 'CHR'] = None

# .iat & .iloc are indices variants of the above
pp.iat[3, 1] = 10.0
pp.iloc[2, 0] = 30

pp

```

