# SO 01-10 {#so-p10}

```{r 'setup', include=FALSE}
source("_common.R")
```

## Test Data

```{r 'R-DF', decorate=TRUE}
df_r <- readr::read_rds("data/R_01_readr.rds")
str(df_r)
```

```{python 'Y-Feather', decorate=TRUE}
df_y = pyarrow.feather.read_feather("data/Y_01_pyarrow.feather")
df_y.info(memory_usage = False)

```

## 001-010

- 001 `r q_so("q5963269")`: How to make a reproducible example or Minimum Working Example (MWE)
  - R: All Libraries by `library()`, Minimal or standard Dataset by `data()` & `dput()`, Minimal Running Code, `set.seed()` for reproducibility, `sessionInfo()` for Environment and OS Information, Revert all the changes made to environment, R Options etc.
  - Python: Use `r q_link("pyarrow.feather.write_feather()")`
    - `r q_link("lib.functions.repr()")` does not keep the `r q_link("pandas.DataFrame.dtypes")`. Similar problem may occur with `r q_link("pandas.DataFrame.to_dict()")` & `r q_link("pandas.DataFrame.to_csv()")`. 
    - `r q_link("pandas.DataFrame.to_pickle()")` & `r q_link("pandas.DataFrame.to_hdf()")` are unsafe.

```{r 'R-data', decorate=TRUE}
data(package = "dplyr")$results[ , "Item"]        #Load or List Datasets
```

```{r 'R-dput', decorate=TRUE}
aa <- df_r
dput(aa)                                          #Write an Object to Console
```

```{python 'Y-repr', decorate=TRUE}
pp = {'x': (11, 22), 'y': [11, 22]}
assert(eval(repr(pp)) == pp)                      #May lose dtypes

```

```{python 'Y-DF-Dict', decorate=TRUE}
pp = df_y.copy()
assert(pd.DataFrame(pp.to_dict()).equals(pp))

pp.to_dict()                                      #May lose dtypes

```



