# Window Functions {#window-p11}

```{r 'setup', include=FALSE}
source('_common.R')
```

## Basics

A window function is a variation on an aggregation function. Where an aggregation function, like `sum()` and `mean()`, takes n inputs and return a single value, a window function returns n values. The output of a window function depends on all its input values, so window functions do not include functions that work element-wise, like `round()`. Window functions include variations on aggregate functions, like `cumsum()` and `dplyr::cummean()`, functions for ranking and ordering, like `rank()`, and functions for taking offsets, like `dplyr::lead()` and `dplyr::lag()`. [(Refer)](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html)

## Ranks

- Python Pandas `r q_link('pd.Series.rank()')` Function & R `{dplyr}` package 
  - NOTE: `r q_link('pd.Series.rank()')` can return cumulative distribution equivalent to `dplyr::cume_dist()` but currently it cannot provide relative rank like `dplyr::percent_rank()`
- `dplyr::row_number()` attributes a unique value to each row
- `dplyr::min_rank()` attributes the same row number to the same value, leaving 'holes'
- `dplyr::dense_rank()` attributes the same row number to the same value, leaving no 'holes'
  - Ex: My salary might be the 2nd-highest (Dense Rank) the company pays. But, I might be the 4th-highest (Rank) paid employee if the same top salary is paid to 3 other people.
- `dplyr::percent_rank()` Rescaling min_rank() to [0, 1]
- `dplyr::cume_dist()` Proportion of all values less than or equal to the current rank.
- `ntile()` a rough rank, which breaks the input vector into n buckets

```{r 'R-Rank', decorate=TRUE}
aa <- c(12, 15, 11, 13, NA, 12)
bb_row <- dplyr::row_number(aa)
bb_row_r <- rank(aa, ties.method = 'first', na.last = 'keep')
bb_row

dd_min <- dplyr::min_rank(aa)                     #Two 12, so skips once for 13
dd_min_r <- rank(aa, ties.method = 'min', na.last = 'keep')
dd_min

dplyr::dense_rank(aa)                             #Consecutive Rank (Dense Rank)

dplyr::percent_rank(aa)                           #Pandas do not have this

dplyr::cume_dist(aa)                              #Matches Pandas Percent Rank

dplyr::ntile(aa, n = 2)

stopifnot(all(identical(bb_row, bb_row_r), identical(dd_min, dd_min_r)))
```

```{python 'Y-Rank', decorate=TRUE}
pp = pd.Series([12, 15, 11, 13, None, 12])
print(*pp)                  #Print Horizontal by using unpacking operator (*)

print(*pp.rank(method = 'first'))                 #Row Number


print(*pp.rank(method = 'min'))                   #Two 12, so skips once for 13


print(*pp.rank(method = 'dense'))                 #Consecutive Rank (Dense Rank)


qq = (pp.rank(method = 'min') - 1) / (pp.count() - 1)       #Percent Rank
print(*qq)


print(*pp.rank(method = 'max', pct = True))       #Cumulative Distance

```

## Data: Lahman

```{r 'R-Tbl', decorate=TRUE}
# DF | Tibble | Subset | Sort | Semi Join |
batting <- Lahman::Batting |> 
    as_tibble() |> 
    select(playerID, yearID, teamID, G, AB, R, H) |> 
    arrange(playerID, yearID, teamID) |> 
    semi_join(Lahman::AwardsPlayers, by = "playerID")

# Grouped Tibble
players <- batting |> group_by(playerID)
```

```{r 'R-Tbl-Data', decorate=TRUE}
# Grouped Players
players |> head()
```

## Test

```{r 'R-Filter', decorate=TRUE}
# For each player, find the two years with most hits
dplyr::filter(players, min_rank(desc(H)) <= 2 & H > 0)


# Within each player, rank each year by the number of games played
dplyr::mutate(players, G_rank = min_rank(G))
```

```{python 'Y-Verify', decorate=TRUE, include=FALSE}
# Count & List the Imported Modules in Python 
q_mods = [v.__name__ for k, v in globals().items() 
    if type(v) is types.ModuleType and not k.startswith('__')]
len(q_mods)
', '.join(q_mods)

```

```{r 'R-Verify', decorate=TRUE, include=FALSE}
if(FALSE) py_config()         #Python Configuration
if(FALSE) q_url[ , 'URL']     #List of URL of this Page
if(FALSE) q_()                #R Objects of this Page excluding 'q_*'
```
